---
title: "Homework1"
author: "Hong Xu"
date: "August 27, 2015"
output: pdf_document
---
1. **Read the rain dataset**
a. read the data and make it a dataframe called rain.df
```{r}
rain.df <- read.table("rnf6080.dat")
```

b. how many rows and columns?
```{r}
print(nrow(rain.df))
print(ncol(rain.df))
```

c. what are the names of the columns?
```{r}
print(colnames(rain.df))
```

d. what is the value of row 5, column 7?
```{r}
print(rain.df[5, 7])
```

e.  Display the second row of rain.df in its entirety.
```{r}
print(rain.df[2,])
```

f. 
```{r}
names(rain.df) <- c("year","month","day",seq(0,23))
head(rain.df)
# what the command does is to assign new column names so that they make more sense.
# now we can see clearly that the first three columns represent the year, month, and day
# and the rest columns are the rainfall level at the specific hour (0-23) of that day 
```

g.Create a new column in the data frame called daily, which is the sum of the rightmost 24
columns. With this column, create a histogram of the values in this column, which are supposed
to be daily rainfall values. What is wrong with this picture?
```{r}
rain.df$daily <- apply(rain.df[, seq(4, 27)], 1, sum)
hist(rain.df$daily)
# problem: there are negative values in the data! - which is impossible for rainfall levels
```

h. Create a new data frame rain.df.fixed that takes the original and fixes it for the apparent flaw you have discovered. Having done this, produce a new histogram with the corrected data and
explain why this is more reasonable.
```{r}
rain.df.fixed <- rain.df[rain.df$daily>0, ]
hist(rain.df.fixed$daily)

# this new plot is more reasonable since it reflects what you expect to see for 
# rainfall levels - majority of them have lower value and as the value increases, 
# the fewer occurances.
```

2. **Syntax and class typing**
a. 
```{r, error=TRUE}
vector1 <- c("5", "12", "7", "32")
max(vector1)
# correct code, erroneous result: the data type of vector is character, 
# so with max you don't get the expected result

sort(vector1)
# correct code, erroneous result: same reason as above

sum(vector1)
# erroneous code: character type does not support add operation 
```
b.
```{r, error=TRUE}
vector2 <- c("5",7,12)
vector2[2] + vector2[3]
# error: c("5",7,12) will coerce 7, 12 to the type of "5", 
# which is character, which cannot be added

dataframe3 <- data.frame(z1="5",z2=7,z3=12)
dataframe3[1,2] + dataframe3[1,3]
# correct results: dataframe support heterogeneous data type 

list4 <- list(z1="6", z2=42, z3="49", z4=126)
list4[[2]]+list4[[4]]
# correct results: list support heterogeneous data type 
# and using [[ will pull out the actual component of the list

list4[2]+list4[4]
# error: using [ will return a list thus cannot be added
```

3. **Working with functions and operators**
a. 
```{r}
seq(1, 10000, by=372)
seq(1, 10000, length.out=50)
```
b.
```{r}
rep(1:3, times=3)
rep(1:3, each=3)
# rep will repeat the whole vector (the first argument x) the "times" times
# while for "each", rep will repeat each element of x that many times, in order
```


